---
title: "HW 1"
output: github_document
date: '2022-09-20'
author: Lily Franks
---
***Setup***
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)

library(nycflights13)
library(ggplot2)
```

```{r}
flights
flights2 = flights
```

***Question 1:***
**How many flights have a missing dep_time?** 
```{r}
sum(is.na(flights$dep_time))
```
**What other variables are missing?**
```{r}
names(which(colSums(is.na(flights))>0))
```
**What might these rows represent?** Cancelled flights. 

***Question 2:***
**Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.**
```{r}
convert_mins = function(val) {
  if (!is.na(val)) {
    hour = as.numeric(substr(val,1,nchar(val)-2)) #the hour
    min = val - round(val, -2) #the min 
    return ((hour*60) + min)
    
  } 
  
  else {
    next
    
  }
  
}
flights$dep_time = convert_mins(flights2$dep_time) 
flights$sched_dep_time = convert_mins(flights2$sched_dep_time) 
flights 
```



***Question 3:***
**Look at the number of canceled flights per day. Is there a pattern? Is the proportion of canceled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with ggplot(aes(x= ,y=)) + geom_point()**

```{r}
flights3 = flights %>%
  mutate(miss_flight = case_when((is.na(dep_time) & is.na(arr_time) & is.na(arr_delay) & is.na(tailnum) & is.na(air_time)) ~ 1, TRUE ~ 0)) %>%
  group_by(day) %>%
  summarise(tot_miss_flight = sum(miss_flight), mean_dely = mean(dep_delay,na.rm=TRUE)) %>%
  ggplot(aes(x = day)) + 
  geom_point(aes(y = tot_miss_flight), color = "red") +
  geom_point(aes(y = mean_dely), color = "blue") +
  ylab("Total Missed Flights (Red) & Delays (Blue)") 

flights3
``` 

**Response:** Flight delays and cancellation do not share a strong correlation and do not seem to be related. 

